Initial Exploration

### 2019-02-24 Text_classifier_models.Rmd
update

---
title: "Text Classification Primer"
author: "Dan Johnson"
date: "1/9/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

For this text classification project I am using an initial codebook and corpus of hand-coded TA comments to test various approaches to semi- and fully automated classification methods and, if appropriate, revise the underlying classification scheme.

This page provides background on general mechanics of text classification and various classifier methods. Other pages describe the dataset structure [LINK] (TA comments corpus), and [LINK] development of the initial codebook.

#Workflow
Text classification problems can be broken down into four overlapping sub-problems. Each piece may be more or less important to the overall problem, but all need to be considered:

@@

Consider a speaker, who uses the term "allow" multiple times throughout the speech, compared to an another speaker who uses terms "allow, concur, acquiesce, accede, and avow"" for the same word. The latter speech has more lexical diversity than the former. Lexical diversity is widely believed to be an important parameter to rate a document in terms of textual richness and effectiveness.

Lexical diversity, in simple terms, is a measurement of the breadth and variety of vocabulary used in a document. The different measures of lexical diversity are TTR, MSTTR, MATTR, C, R, CTTR, U, S, K, Maas, HD-D, MTLD, and MTLD-MA.

@@

Understanding how chatbots work can help us understand text classifiers, because the classifier is the fundamental piece that makes the chatbot work. Let’s look at the inner workings of a Multinomial Naive Bayes algorithm for text classification and natural language processing (NLP).

@@

###Pattern Matchers
Early chatbots used **pattern matching** to classify text and produce a response. This is often referred to as “brute force” as the author of the system needs to describe every pattern for which there is a response.

@@

A simple pattern matching definition:
```
<aiml version = "1.0.1" encoding = "UTF-8"?>
   <category>
      <pattern> WHO IS ALBERT EINSTEIN </pattern>
      <template>Albert Einstein was a German physicist.</template>
   </category>

@@

   <category>
      <pattern>DO YOU KNOW WHO * IS</pattern>
      <template>
         <srai>WHO IS <star/></srai>
      </template>
   </category>
</aiml>
```

@@

###Algorithm-Based Classification
Creating a brute-force pattern matcher is daunting: for each unique input a pattern must be available to specify a response. An algorithm-based method approaches the problem mathematically.

@@

Look at this sample training set:
```
class: weather
    "is it nice outside?"
    "how is it outside?"
    "is the weather nice?"

class: greeting
    "how are you?"
    "hello there"
    "how is it going?"
```

Let’s classify a few sample input sentences:
```
input: "Hi there"
 term: "hi" (no matches in either class)
 term: "there" (matches a word in class: greeting)
 classification: greeting (score=1)

input: "What’s it like outside?"
 term: "it" (matches words in class: weather (2), greeting(1))
 term: "outside (class: weather (2) )
 classification: weather (score=4)
```

@@

Many text and spoken word frameworks use similar algorithms such as this to classify intent. Most of what’s taking place is word counting against training datasets. It’s "naive" but surprisingly effective.

@@

Algorithms also can be separated into un-supervised and supervised methods.

**Unsupervised learning** uses test data that has not been labeled, classified or categorized. Unsupervised learning identifies commonalities in the training data, then classifies new texts based on the presence or absence of such commonalities.

@@

***
***

>>>>

---
title: "Text Classification Primer"
date: "2/21/2019"
bibliography: library.bib
biblio-style: apalike
---

```{r, child="_setup.Rmd"}
```

#Overview
This page provides general background on theory and approaches to text classification, and descriptions of several classification methods.

####\

#Basic Workflow
Text classification problems can be broken down into four overlapping sub-problems. Each piece may be more or less important to the overall problem, but all need to be considered:

@@

>Consider a speaker, who uses the term "allow" multiple times throughout the speech, compared to an another speaker who uses terms "allow, concur, acquiesce, accede, and avow"" for the same word. The latter speech has more lexical diversity than the former. Lexical diversity is widely believed to be an important parameter to rate a document in terms of textual richness and effectiveness.

>Lexical diversity, in simple terms, is a measurement of the breadth and variety of vocabulary used in a document. The different measures of lexical diversity are TTR, MSTTR, MATTR, C, R, CTTR, U, S, K, Maas, HD-D, MTLD, and MTLD-MA.

@@

```{}
Understanding how chatbots work can help us understand text classifiers, because the classifier is the fundamental piece that makes the chatbot work. Let’s look at the inner workings of a Multinomial Naive Bayes algorithm for text classification and natural language processing (NLP).

@@

Pattern Matchers
Early chatbots used **pattern matching** to classify text and produce a response. This is often referred to as “brute force” as the author of the system needs to describe every pattern for which there is a response.

@@

A simple pattern matching definition:

<aiml version = "1.0.1" encoding = "UTF-8"?>
   <category>
      <pattern> WHO IS ALBERT EINSTEIN </pattern>
      <template>Albert Einstein was a German physicist.</template>
   </category>

@@

   <category>
      <pattern>DO YOU KNOW WHO * IS</pattern>
      <template>
         <srai>WHO IS <star/></srai>
      </template>
   </category>
</aiml>

@@

Algorithm-Based Classification

Creating a brute-force pattern matcher is daunting: for each unique input a pattern must be available to specify a response. An algorithm-based method approaches the problem mathematically.

@@

Look at this sample training set:

class: weather
    "is it nice outside?"
    "how is it outside?"
    "is the weather nice?"

class: greeting
    "how are you?"
    "hello there"
    "how is it going?"

Let’s classify a few sample input sentences:

input: "Hi there"
 term: "hi" (no matches in either class)
 term: "there" (matches a word in class: greeting)
 classification: greeting (score=1)

input: "What’s it like outside?"
 term: "it" (matches words in class: weather (2), greeting(1))
 term: "outside (class: weather (2) )
 classification: weather (score=4)

@@

Many text and spoken word frameworks use similar algorithms such as this to classify intent. Most of what’s taking place is word counting against training datasets. It’s "naive" but surprisingly effective.
```

@@

Algorithms also can be separated into un-supervised and supervised methods. 
**Unsupervised learning** uses test data that has not been labeled, classified or categorized. Unsupervised learning identifies commonalities in the training data, then classifies new texts based on the presence or absence of such commonalities.

@@

####\

### 2019-02-24 build.Rmd
update

---
title: "The Build"
author: "Dan Johnson"
date: "2/22/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Overview
This page links to sub-pages that describe how various components used in the project were built. Follow the links for details.

Extracting TAs' comments: https://adanieljohnson.github.io/default_website/docxcommentextraction.html

@@

Current Naive Bayes classifier: https://adanieljohnson.github.io/default_website/NB_2r2c.html

@@

#Describe an Element

#Describe an Element

>>>>

---
title: "The Build"
date: "2/22/2019"
bibliography: library.bib
biblio-style: apalike
---

```{r, child="_setup.Rmd"}
```

#Overview
This page links to sub-pages that describe the build process for various components used in the overall project. Follow the links for details.

Extracting TAs' comments (*draft*): https://adanieljohnson.github.io/default_website/docxcommentextraction.html

Codebook for classifying TAs' comments: https://adanieljohnson.github.io/default_website/codebook.html

@@

Current version of the Naive Bayes classifier: https://adanieljohnson.github.io/default_website/NB_2r2c.html

Expanding the Naive Bayes classifier to an Ensemble Model (*draft*): https://adanieljohnson.github.io/default_website/Ensemble_Notes.html

@@

### 2019-02-24 codebook.rmd
update

---
title: "Comments Codebook"
date: "2/23/2019"
bibliography: library.bib
biblio-style: apalike
---

@@

We evaluating TAs’ comments on student reports to see if and how their comments change as students get more feedback from automated system.

To build the data table, commented reports from a single semester (MS Word .docx format) were processed using BaSH scripts to extract relevant file data (author, ID#, etc.) and individual TA comments from the raw XML, and write the data to a single CSV file. TA comments containing multiple points or suggestions were replicated and edited to produce a working data table with only one comment per entry row.

Inclusion and exclusion criteria for assigning TA comments to categories and sub-categories was established iteratively using two training sets of 120 and 1200 comments selected randomly from the full dataset of ~11,000. Each main category (Subject, Structure, etc.) was divided into 4-8 loosely defined provisional sub-categories.Each comment from the 120-item training set was assigned to a provisional sub-category within each category. Sub-categories from different main categories were not mutually exclusive; for example, if a comment was assigned to the “Writing Quality” sub-category under the “Subject” category, that comment could be assigned to any of the provisional sub-categories in “Structure,” any of the sub-categories in “Locus,” etc.

@@

Ah, so it didn’t move directly towards the Physarum. Here it’s important to make that distinction in how you took measurements in your methods section. Did you go based on leading edge? If so, this data point would probably be inconclusive in the results section. Did you go based on a central axis upon which you took measurements? Instead, that’d be different.

>>>>

---
title: "The TA Comments Codebook"
date: "2/23/2019"
bibliography: library.bib
biblio-style: apalike
---

@@

We classify TAs’ comments on student reports to see if and how their comments change as students get more feedback from automated system.

To build a TA comments data table, commented reports from a single semester (MS Word .docx format) are processed using R or BaSH scripts to extract relevant file data (author, ID#, etc.) and individual TA comments from the raw XML, and write the data to a single CSV file. TA comments containing multiple points or suggestions were replicated and edited to produce a working data table with only one comment per entry row.

Inclusion and exclusion criteria for assigning TA comments to categories and sub-categories were established iteratively using two training sets of 120 and 1200 comments selected randomly from the full dataset of ~11,000. Each main category (Subject, Structure, etc.) was divided into 4-8 loosely defined provisional sub-categories.Each comment from the 120-item training set was assigned to a provisional sub-category within each category.

Note: sub-categories from different main categories are not mutually exclusive; for example, if a comment is assigned to the “Writing Quality” sub-category under the “Subject” category, that comment can be assigned to any of the provisional sub-categories in “Structure,” any of the sub-categories in “Locus,” etc.

@@

>Ah, so it didn’t move directly towards the Physarum. Here it’s important to make that distinction in how you took measurements in your methods section. Did you go based on leading edge? If so, this data point would probably be inconclusive in the results section. Did you go based on a central axis upon which you took measurements? Instead, that’d be different.

### 2019-02-24 docxcommentextraction.Rmd
update

#Pseudo-Code and Sources
Create a list of files to cycle through OR put all files to process in ONE folder.
For each file, use 'read_docx', then 'docx_extract_all_cmnts'
Use 'do.call("rbind", list(dataframes to merge))' to cycle through and append the extracted comments to ONE dataframe

@@

https://www.rdocumentation.org/packages/analogue/versions/0.17-1/topics/join
Load required Libraries
```{r}
library(docxtractr)
```

```{r}
Commentframe1<-read_docx("/Users/danjohnson/Dropbox/Coding_Tools/R_Environment/R_Projects/DOCX_comment_extraction/TestingReport.docx", track_changes = NULL)

Commentframe2<-read_docx("/Users/danjohnson/Dropbox/Coding_Tools/R_Environment/R_Projects/DOCX_comment_extraction/TestingCommentExtraction.docx", track_changes = NULL)

comments1<-docx_extract_all_cmnts(Commentframe1, include_text = FALSE)
comments2<-docx_extract_all_cmnts(Commentframe2, include_text = FALSE)

#Merging more than two dataframes
#To merge more than two dataframes within R, we can combine our rbind call with do.call.
# > do.call("rbind", list(dataframes to merge))
#
#How this works:
#  The 'do.call' function allows us to use any existing R functions, such as 'rbind', but instead of passing just two dataframes in this instance, we are able to pass multiple.
#
#Therefore, when used alongside rbind, we can use do.call to merge more than two dataframes at once.
#
#Putting this all together, we can merge our dataframes back together as follows:
# > ufoMerged <- do.call("rbind", list(ufo1, ufo2, ufo3, ufo4))
#
#We can confirm our data has been successfully merged by checking the number of rows:
#
#> NROW(ufoMerged)
#[1] 80332

#comments1And2 <- rbind(comments1,comments2)
do.call("rbind", list(comments1,comments2))

write.csv2(comments1And2, file = "TAcomments.csv")
```

>>>>

The testing dataset used in this project was created using BaSH scripting and one-off tools. This page documents progress on a workflow to replicate that process using R tools.

####\

#General Approach
Create a list of files to cycle through OR put all files to process in ONE folder. For each file, use 'read_docx', then 'docx_extract_all_cmnts'. Use 'do.call("rbind", list(dataframes to merge))' to cycle through and append the extracted comments to ONE dataframe

####\

#Code Sources, Examples

@@

https://www.rdocumentation.org/packages/analogue/versions/0.17-1/topics/join
Load required Libraries

####\

#Sample Working Code
```{r}
library(docxtractr)
```

Call in Word files using ```read_docx```. Use ```docx_extract_all_cmts``` to pull out comments and ```rbind``` to combine them.

To merge more than two dataframes within R, combine an ```rbind``` call with ```do.call```. The 'do.call' function uses any existing R functions, such as 'rbind', but instead of passing just two dataframes, it passes multiple data frames. Then use ```write.csv2``` to write the data to a CSV format.

```{}
Commentframe1<-read_docx("/Users/danjohnson/Dropbox/Coding_Tools/R_Environment/R_Projects/DOCX_comment_extraction/TestingReport.docx", track_changes = NULL)
comments1<-docx_extract_all_cmnts(Commentframe1, include_text = FALSE)

Commentframe2<-read_docx("/Users/danjohnson/Dropbox/Coding_Tools/R_Environment/R_Projects/DOCX_comment_extraction/TestingCommentExtraction.docx", track_changes = NULL)
comments2<-docx_extract_all_cmnts(Commentframe2, include_text = FALSE)

# comments1And2 <- rbind(comments1,comments2)
# OR
# comments1And2 <- do.call("rbind", list(comments1,comments2))

#Check data have been successfully merged by checking the number of rows
NROW(comments1And2)
# [1] 32

#When all lines are appended, write the CSV file 
write.csv2(comments1And2, file = "TAcomments.csv")

```

### 2019-02-24 initial_exploration.Rmd
update

---
title: "Initial Exploration"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Overview
The first question to answer before making deeper comparisons is, **are word frequencies different between TAs' comments from different sub-categories?** To that end, I ran three different analyses of word usage frequency.

*  __Pearson correlations for word use frequencies__. This only compares words found in all of the Subject sub-categories
*  __Word frequency differences based on log-odds ratios__. This highlights terms that are the most different between sub-categories.
*  __N-grams mapping__. This identifies text strings, phrases, and terms that are unique (or nearly unique) to one sub-category.

@@

#N-Grams Mapping 
##Are Certain Terms and Phrases Unique to One Sub-Category?
One of our earliest observations was that the n-gram "basic criteria" only appeared in TA comments from three different Subject sub-categories. This suggests certain terms and n-grams might be eliminated

However it w  the relative frequency was much higher
```{}
#Pseudo-code
Import comments
Tokenize the comments as 1-, 2, 3-grams
Add counts within subset, calculate frequency
Split by subcategory
Align 4 subcategories, highest to lowest

>>>>

---
title: "Initial Exploration"
date: "2/21/2019"
bibliography: library.bib
biblio-style: apalike
---

```{r, child="_setup.Rmd"}
```

#Overview
The first question to answer before making deeper comparisons is, **are word frequencies different between TAs' comments from different sub-categories?** This page describes the methods used for this initial exploration.

Results of these analyses are discussed on this page: https://adanieljohnson.github.io/default_website/findings.html

#The Analyses
I ran three different analyses of word usage frequency.

*  __Pearson correlations for word use frequencies__. This only compares words found in all of the Subject sub-categories
*  __Word frequency differences based on log-odds ratios__. This highlights terms that are the most different between sub-categories.
*  __N-grams mapping__ (*draft*). This identifies text strings, phrases, and terms that are unique (or nearly unique) to one sub-category.

@@

#N-Grams Mapping 
##Are Certain Terms and Phrases Unique to One Sub-Category?
One of our earliest observations was that the n-gram "basic criteria" only appeared in TA comments from three different Subject sub-categories. Of these, the phrase made up <0.5% of comments in two categories, but in the third category, "basic criteria" made up >5% of all comments. This suggests that the "basic criteria" n-gram is essentially diagnostic for that sub-category. This suggests other terms and n-grams that are eliminated in the previous two methods may be equally diagnostic for one sub-category.

The block below describes a pseudo-code approach.
```{}
Import comments
Tokenize the comments as 1-, 2, 3-grams
Add counts within subset, calculate frequency
Split by subcategory
Align 4 subcategories, highest to lowest
